Corotating Dust Clumps Around Adolescent Low-Mass Stars: Four Years of TESS

### 2023-06-28 paper/ms.tex
update ms

\title{Corotating Dust Clumps Around Adolescent M-Dwarfs: Four Years of TESS}

@@

% 250 word max!
\begin{abstract}
  Complex quasiperiodic variables (CQVs) are low-mass
  pre-main-sequence stars with nearly periodic optical modulation that
  is likely induced by long-lived transiting dust clumps in orbit at
  the Keplerian corotation radius.  Here, we report new CQVs
  discovered in TESS data collected between July~2018 and Sep~2022.
  From our analysis of \nstarssearched\ K- and M-dwarfs with $T$$<$16
  and $d$$<$150\,pc, we find \ngoods\ gold-standard CQVs.  Most of
  these discoveries are new, and they include the brightest and
  closest examples of this class of object known ($T$$\approx$11;
  $d$$\approx$20\,pc).  A few objects are outliers among outliers; LP
  12-502 for instance shows a ``dip complex'' with a period and total
  duration that are fixed over more than 1000 cycles, but in detail
  this object shows anywhere from four to eight local minima per
  cycle.  LP-502 also demonstrated drastic shape changes over less
  than one cycle, and at times displayed distinct superposed periods
  simultaneously.  Generally speaking, we find that none of the CQVs
  maintain a fixed light curve shape over timescales of more than a
  few hundred cycles -- they are all {\it quasi}periodic.  We revisit
  the arguments for why corotating dust clumps are the most likely
  explanation, though there are other possibilities, and all lines of
  evidence are currently indirect.  In the future, we expect that our
  sample will help facilitate modeling and observational efforts aimed
  at understanding these objects, and connecting them to the broader
  contexts of star, disk, and exoplanet evolution.
\end{abstract}

@@

The mosaic in Figure~\ref{fig:singlemosaic} shows phased light curves
for the \ngoods\ bonafide CQVs.  
The objects are sorted first in order of the number of TESS sectors in
which they clearly demonstrated the CQV phenomenon, and secondarily by
descending brightness.
The objects at the top therefore have the most data.

Generally, Figure~\ref{fig:singlemosaic} shows a broad range of
variability, with anywhere from two to eight local minima per cycle.  
Some stars show relatively ordinary modulation during one
continuous portion of the phased light curve, and highly structured
modulation in the remainder of the cycle (e.g. TIC~206544316,
TIC~224283342, TIC~402980664).  Others show structured modulation over
the entire span of a cycle (e.g. TIC~2234692, TIC~401789285,
TIC~425933644, TIC~142173958).  Others still show some mix between
these two modes.

A few objets at first glance seem reminiscent of eclipsing binaries, for
instance TIC~193831684.
In every such case,

>>>>

\title{Corotating Dust Clumps Around Adolescent Low-Mass Stars: Four Years of TESS}

@@

% 250 word max!
\begin{abstract}
  Complex quasiperiodic variables (CQVs) are low-mass
  pre-main-sequence stars with nearly periodic optical modulation.
  The modulation is likely induced by dust clumps in orbit at the
  Keplerian corotation radius.  Here, we report new CQVs discovered in
  TESS data collected between July~2018 and Sep~2022, and analyze
  their variability.  Our analysis of \nstarssearched\ K- and
  M-dwarfs with $T$$<$16 and $d$$<$150\,pc yielded \ngoods\
  gold-standard CQVs.  Most of these discoveries are new, and they
  include the brightest and closest examples of this class of object
  known ($T$$\approx$11; $d$$\approx$20\,pc), as well as the most
  massive ($\approx$0.65\,M$_\odot$).  A few objects are outliers
  among outliers; LP 12-502 for instance shows a ``dip complex'' with
  a period and total duration that are fixed over more than 1000
  cycles, but which in detail shows anywhere from four to eight
  local minima per cycle.  LP-502 also demonstrated drastic shape
  changes over less than one cycle, and at times displayed distinct
  superposed periods simultaneously.  Broadly speaking, we find that
  none of the CQVs maintain a fixed light curve shape over timescales
  of more than a few hundred cycles -- they are all {\it
  quasi}periodic.  We revisit the arguments for why transient
  corotating dust clumps are the most likely explanation, though there
  are other possibilities.  In the future, we expect that our sample
  will help facilitate modeling and observational efforts aimed at
  understanding these objects, and connecting them to the broader
  contexts of star, disk, and exoplanet evolution.
\end{abstract}

@@

The mosaic in Figure~\ref{fig:singlemosaic} shows phased light curves
for the \ngoods\ bonafide CQVs.  The objects are sorted first in order
of the number of TESS 120-second cadence sectors in which they clearly
demonstrated the CQV phenomenon, and secondarily by descending
brightness.  The objects at the top therefore have the most 120-second
cadence data.  The top five objects by this metric are TIC~300651846
($T$$=$13.5, 12 sectors); TIC~402980664 ($T$$=$11.1, 7 sectors);
TIC~89463560 ($T$$=$13.5, 5 sectors); TIC~363963079 ($T$$=$12.9, 5
sectors); and TIC~294328887 ($T$$=$14.2, 4 sectors).  The brightest
five CQVs span 9.3$<$$T$$<$11.1; the faintest five span
14.5$<$$T$$<$15.0.  The fastest five have periods spanning
3.6\,hr$<$$P$$<$6.2\,hr, while the slowest five span
27\,hr$<$$P$$<$38\,hr.

In terms of the light curve shapes, Figure~\ref{fig:singlemosaic}
shows a broad range of variability, with anywhere from two to eight
local minima per cycle.  Some stars show relatively ordinary
modulation during one continuous portion of the phased light curve,
and highly structured modulation in the remainder of the cycle (e.g.
TIC~206544316, TIC~224283342, TIC~402980664).  Others show structured
modulation over the entire span of a cycle (e.g. TIC~2234692,
TIC~401789285, TIC~425933644, TIC~142173958).  Others still show some
mix between these two modes.

A small number of objects at first glance seem reminiscent of
eclipsing binaries, for instance TIC~193831684 or TIC~5714469.  In the
few such cases, we believe that that are unlikely to be eclipsing
binaries due to additional coherent variations in the light curves
that are distinct from any binary phenomenology of which we are aware.
